Privacy-Respecting Type Error Telemetry at Scale

### 2023-09-30 programming23/paper.tex
macros for ID (\uid) and IDE (\IDE)

% TODO BEFORE SUBMISSION
% be consistent between "IDs" and "ids" [SK prefers "IDs" to avoid it being read as a word]
% even more so between "ID" and "id" [NB: "id" is a word in English]
% \usepackage[sort]{natbib} or otherwise manually fix cites to be in ascending order
% redo fig 5 to have better resolution

% TODO BEFORE FINAL
% - uncomment acks

@@

\title{Type Error Telemetry at Scale}
% lean, lite, modest, non-intrusive, private, ...
% {Millions of Type Errors} = 72M forced strict, 1/2M type error is disappointing!
%% \subtitle{Impersonal Telemetry to Measure User Experience}

@@

\newcommand{\code}[1]{\texttt{#1}}
\newcommand{\FILL}{\textbf{FILL}}
\newcommand{\dotscale}[1]{\scalebox{0.72}{#1}}
\newcommand{\wideas}[2]{\makebox[\widthof{#2}][l]{#1}}
\newcommand{\twoline}[2]{\parbox[s]{1.4cm}{\flushleft#1\newline#2}}
\newcommand{\chkYes}{\dotscale{\CIRCLE}}
\newcommand{\chkMaybe}{\wideas{\dotscale{\Circle}}{\chkYes}}
\newcommand{\chkNo}{\wideas{}{\chkYes}}
\newcommand{\pct}[1]{\SI{#1}{\percent}}
\newcommand{\percentile}[1]{{#1}th~percentile}
\newcommand{\ptile}[1]{\textsc{p}{#1}}
\newcommand{\modefont}[1]{\texttt{#1}}
\newcommand{\mnocheck}{\modefont{nocheck}}
\newcommand{\mnonstrict}{\modefont{nonstrict}}
\newcommand{\mstrict}{\modefont{strict}}
\newcommand{\FS}{background}
\newcommand{\zerowidth}[1]{\makebox[0pt][l]{#1}}
\newcommand{\zeroheight}[1]{\raisebox{0pt}[0pt][0pt]{\(#1\)}}
\newcommand{\nK}[1]{#1\textsc{k}}
\newcommand{\stddev}[1]{\nK{#1}}
\newcommand{\gcell}[1]{\cellcolor{green!20}#1}
\newcommand{\ycell}[1]{\cellcolor{yellow!18}#1}
\newcommand{\ocell}[1]{\cellcolor{orange!29}#1}
\newcommand{\rcell}[1]{\cellcolor{red!30}$\!\!$#1$\!\!$}
\newcommand{\gbox}[1]{\colorbox{green!20}{#1}}
\newcommand{\ybox}[1]{\colorbox{yellow!18}{#1}}
\newcommand{\obox}[1]{\colorbox{orange!29}{#1}}
\newcommand{\rbox}[1]{\colorbox{red!30}{#1}}
\newcommand{\tefsfont}[1]{\textsc{#1}}
\newcommand{\tekey}{\tefsfont{te}}
\newcommand{\fskey}{\tefsfont{bg}}
\newcommand{\QALAN}{\FILL{} Alan:}
\newcommand{\panon}{pseudonymized}
\newcommand{\symbox}[1]{{\makebox[0.8em][c]{{#1}}}}
\newcommand{\addsym}{\symbox{\(\uparrow\)}}
\newcommand{\keepsym}{\symbox{\(=\)}}
\newcommand{\dropsym}{\symbox{\(\downarrow\)}}
% \newcommand{\dsymbox}[1]{\makebox[0.8em][c]{#1}}
% \newcommand{\dropsym}{\dsymbox{\raisebox{0ex}[0.9em]{\raisebox{0.5ex}{\vrule width 0.6em height 0.6pt}}}}

@@

  \paragraph{Inquiry}
  %% What problem or question does the paper
  %% address? How has this problem or question been
  %% addressed by others (if at all)?
  Since Luau is currently being used by many creators, we want to collect their feedback
  to improve the language and, in particular, the type system.
  The standard way to collect feedback at a large scale from creators working on
  real projects is to deploy client-side telemetry; however, we cannot
  scrape personal data or proprietary information, which means we cannot
  collect source code fragments, error messages, or even filepaths.
  The research questions are thus about how to conduct telemetry that
  is not invasive and obtain insights from it about type errors.

  \paragraph{Approach}
  %% What was done that unveiled new knowledge?
  We designed and implemented a pseudonymized, randomized telemetry system for Luau.
  Telemetry records include a timestamp, a session id, a reason for sending,
  and a numeric summary of the most recent type analyses.
  This information lets us study type errors over time
  without revealing private data.
  We deployed the system in Roblox Studio during Spring 2023
  and collected over 1.5 million telemetry records from over 340,000 sessions.

@@

Creators program using the
{Luau} programming language~\cite{luau-lang.org},
an extension of {Lua~5.1~\cite{lua}}.
The main addition in Luau is a static type system that infers
types for all Luau programs on the fly, as creators modify the code.
These types are used primarily in IDE tooling such as autocomplete and
in API documentation~\cite{luau-autocomplete}, and indeed, creators
may be unaware that a typechecker is analyzing their code.
However, creators can opt in to receiving type error reports and they can write
their own types to guide designs and document their intentions.

Due to the broad community of creators, the goals of the Luau type
system are rather unique~\cite{bfj-hatra-2021}.
Whereas a traditional type system focuses on compilation and memory
safety, Luau takes a lenient approach by default and lets creators
gradually~\cite{st-sfp-2006,tf-popl-2008} migrate to rigorous checks
one module at a time.
In this way, for example,
students in code camps may begin in the \mnocheck{} mode that reports only
syntax errors; hobbyists can use \mnonstrict{} mode to identify
would-be runtime errors; and developers who need to ensure software
quality can rely on \mstrict{} mode.
(No matter the mode, a {\FS{}} type analysis still drives the IDE tools.)
Luau's types aim to support untyped designs, in the spirit of migratory
typing~\cite{tfffgksst-snapl-2017}, so that creators can switch modes without
needing to restructure their code in major, potentially-breaking, ways.

In this paper, we investigate methods for measuring the effectiveness
of the {Luau} type system.
The goal is to collect feedback at a large scale, with thousands
of participants maintaining real codebases.
Consequently, the measurements cannot reveal \emph{any} information about
source code, as it may contain personal data, proprietary algorithms, novel
game designs, and so on.
In comparison to prior work~(\cref{s:related}), which
with few exceptions~(e.g.,\cite{hlzbr-ecoop-2021}) is small in scale
or collects source code, we performed a large-scale study using \panon{}
{telemetry}.

@@

\paragraph{Contributions}
\begin{itemize}
  \item
    Design of a low-overhead telemetry system
    that reports on type errors without revealing source code.

@@

{Roblox Studio} is a workbench that
combines {3D creation} tools and an Integrated
Developer Environment (IDE), as seen in \cref{fig:roblox-studio}.
The IDE includes an optional \emph{Script Analysis} widget that
reports syntax errors, type errors, and problems identified by
lint tools. The main editor widget can also highlight
source locations in code where reported errors occur.

\paragraph{Type Analysis Modes, \FS{} Analysis}

@@

In addition to the main type analysis, Roblox Studio takes a second pass over
every codebase with a \emph{\FS{}} analysis to infer autocomplete suggestions
and drive other IDE tools.
To infer precise types,
the \FS{} analysis uses rigorous checks very similar to those of
\mstrict{} mode (but not identical: see \cref{s:strict-vs-forcedstrict})
no matter what mode the creator declared for the script.
(Internally to Roblox Studio, this analysis is called \emph{forced strict}
because it ignores the creator's choice and applies strong checks.)
When it fails to infer a type, \FS{} analysis defaults to the unknown gradual
type~\cite{st-sfp-2006}.
Creators cannot see the errors reported by \FS{} analysis.

@@

\begin{figure}[t]\centering
  \includegraphics[width=.35\textwidth]{img/roblox-studio.png}
  \hspace{1cm}
  \includegraphics[width=.35\textwidth]{img/roblox-studio-ide.png}
  \caption{{Roblox Studio 3D creation} tools (left) and IDE (right).}
  \label{fig:roblox-studio}
\end{figure}

@@

Scripts come in two flavors: \emph{module scripts} and \emph{non-module
scripts}.
(Though, throughout this paper we use the word ``module''
to refer to arbitrary scripts.)
Module scripts provide reusable libraries, which may be
{imported} by other scripts.
Module scripts therefore form a graph in general,
though it is an error in \mstrict{} mode to create an import cycle
and edges are removed to make the graph acyclic.
%% \QALAN{}: do all modes remove the edges?

@@

Roblox Studio has a framework for reporting telemetry.
Various features in the IDE use telemetry to measure their effectiveness.
The Luau open source toolchain does \emph{not} report telemetry,
as it is designed to be used in build environments such as Continuous Integration
servers, where hermetic deterministic builds are expected.

We added a telemetry subsystem to Roblox Studio in order to study type errors
in Luau code.
This section explains the constraints our telemetry had to meet~(\cref{s:tel-limitations})
and the final design~(\cref{s:telemetry-records}).

@@

Telemetry in Roblox Studio faces three major constraints:
\begin{itemize}

@@

\end{itemize}
Cox~\cite{transparent-telemetry} discusses in detail the privacy implications
and tradeoffs in telemetry.

In addition, due to the architecture of Roblox Studio
there is some information that is not available to the
typechecker:
\begin{itemize}
  \item
    Lifecycle events including as save, run, quit, and publish.\footnote{Publish events
    cannot be reported at any rate because they are public, and could be used
    to match a telemetry session to a creator.}
  \item
    GUI state, in particular, whether the Script Analysis widget is visible.
\end{itemize}
Our type error telemetry has to meet all of the above constraints.

@@

\begin{figure}[t]\centering
  \begin{tabular}{|l|l|l|l|l|l|l|}\hline
    ID & Time & Mode & Reason & Size Info & Overall Counts & Edit Range Counts
    \\\toprule
    \multicolumn{6}{c}{\raisebox{.5\normalbaselineskip}[0pt][0pt]{$\underbrace{\hspace{9cm}}$}} &
    \multicolumn{1}{c}{\raisebox{.5\normalbaselineskip}[0pt][0pt]{$\underbrace{\hspace{3cm}}$}} \\[-0.5ex]
    \multicolumn{6}{c}{fixed-size} & \multicolumn{1}{c}{variable-size}
  \end{tabular}
  \caption{Structure of a telemetry record. There are 13 overall counts and up to 70 edit range counts in each record.}
  \label{f:tdata}
\end{figure}

The type error telemetry for a Roblox Studio session is gathered as a series of
records, all with the same \panon{} session identifier.
This allows us to correlate telemetry across a single
session, but not between sessions.
In order to avoid swamping the telemetry servers, the client reports on a subset
of telemetry events via two levels of uniform random sampling:
\begin{itemize}
  \item
    1\% of Roblox Studio sessions generate \emph{any} type error telemetry, and
  \item
    0.5\% of all type analyses (approximately 1 in 200 keystrokes)
    in an enrolled session generate a telemetry record.
\end{itemize}
In addition to sending telemetry for randomly-selected keystrokes,
our system sends a telemetry record every time the session changes
focus from one module to another.
These records provide important context for mode switches:
if one record uses \mstrict{} mode but the next uses \mnocheck{},
then it is critical to know whether the creator downgraded modes
or simply switched to another module.

\Cref{f:tdata} shows the structure of our telemetry records:
\begin{enumerate}
  \item
    ID: a \panon{} identifier for the current session.
  \item
    Time: one timestamp from the client and one from the server.
    %% timestamp explanation \cref{s:data-cleaning}
  \item
    Mode: type analysis mode of the current file: \mnocheck{},
    \mnonstrict{}, or \mstrict{}.
  \item
    Reason: a flag that explains whether this record was sent due to a
    randomly-selected keystroke or a module switch.
  \item
    Size Info: number of lines in the codebase and number of lines in the edit range.
  \item
    Overall Counts: summarizes type errors and \FS{} errors during the
    last two invocations of type analysis, and additionally reports the number
    of times type analysis hit an internal limit on problem
    size~(see~\cref{s:code-too-complex} for details).
    For each invocation and each kind of
    analysis error, there are three summary counts:
    %% grand total = 12
    \begin{itemize}
      \item total number of errors,
      \item errors in the current module, and
      \item errors in the current edit range.
    \end{itemize}
  \item
    Edit Range Counts: these list specific errors that arose in the last two
    type analyses and overlap with the current edit range.
    Since there are 35 possible errors~(\cref{s:context}), there can be up to 70 counts
    in a telemetry record.
    Exactly how to interpret this data depends on which analysis it came from:
    \begin{itemize}
      \item
        Errors in the latest type analysis are currently visible to the
        creator.
        These may have been introduced by the changes in the edit range.
      \item
        Errors from the previous type analysis were visible before the creator
        made the latest edits.
        These may have motivated some of the changes in the edit range.
    \end{itemize}
\end{enumerate}

@@

This telemetry design is admittedly coarse-grained.
For instance, it does not distinguish between edits that ignore a type
error from edits that remove one error while introducing another.
We acknowledge this and other threats in~\cref{s:threats}.
The main advantages of this telemetry are the small size of each record
and complete the lack of private information.
Furthermore, despite the limitations, this telemetry supports a variety of
interesting conclusions that we showcase in the next section.

@@

\Cref{f:records-per-hour} provides a time-ordered distribution of the data.
The x-axis is labeled with a California time zone; the thin vertical lines
mark the start of a new day.
Each of the blue bars in the plot bar is a bucket that counts the number of
records generated per hour according to a client-side timestamp.
Buckets contain a few dozen to over 3,000 records, typically about 1,000.
%% \FILL{} investigate one tower, what's happening?? all one session?
Telemetry ended in early April; the long tail is due to developers
updating Roblox Studio later.
% BEN: I see no basis for this; nor is this at all interesting.
%% ben: basis = it's where the peaks are
% We do not know which timezone the records originated in, but since the counts
% tend to peak midday California time it seems that the majority
% of Roblox developers are following a Western Hemisphere schedule.
The regular pattern of the peaks suggests that many creators
follow a regular schedule; these may be
professional developers.
The tall weekend (shaded) regions could be because
Roblox has a significant school-aged creator community.

@@

A close inspection of the data revealed two anomalies:
\begin{enumerate}
  \item
    Some records have identical timestamps and session IDs.
    We keep the first such record and discard the rest.
    (A human would have to switch modules twice or enter several keystrokes
     within 1ms to generate two such records legitimately.)

@@

  {Among the mixed-mode sessions:
        176 contain a mode upgrade,
        233 contain a mode downgrade, and
        512 contain modules with different modes.}

@@

%% mostly pure, rare mixes
Most sessions (\pct{99.82}) stick to a single analysis mode.
They never change the mode in the current module and never
switch focus to a module with a different analysis mode.
Among the other, mixed-mode sessions, most of those (\pct{80}) switch to a
module with a different mode, about half contain at least one edit that
upgrades to a more strict mode, and about half contain a downgrade to a
less-strict mode.
There are 263 total upgrades (across all sessions) and 320 total downgrades.

@@

\subsection{Type Error Popularity}
\label{s:type-error-count}

@@

\Cref{t:type-error-count} sorts type errors by their popularity
in \mnonstrict{} and \mstrict{} mode.
The purpose of this table is to discover and rank all errors that
appear in the dataset, thus it is based on both keystroke and
module-switch telemetry records even though edit ranges are not well-defined
for the latter.
The table does not include \mnocheck{} because all its type errors are
\code{SyntaxError}s by definition.

@@

In other professional-grade IDEs, such as VSCode~\cite{vscode-telemetry},
IntelliJ~\cite{intellij-telemetry}, and \code{.NET}~\cite{dotnet-telemetry}, telemetry may include all sorts of
data: error reports, source code fragments, time\-stamps, and filepaths.
Users may be able to opt out of some telemetry, but the details depend on the
license agreement.
Furthermore, at least in VSCode, IDE extensions can report their own telemetry.
While this data can be invaluable for discovering bugs in production, it
must be handled with extreme care.

The Blackbox project similarly collects telemetry in broad strokes.
Every Java program submitted for compilation goes into the Blackbox database (minus
any comments above the main class) along with the compiler's response~\cite{bkmu-sigcse-2014}.
%% Though somewhat intrusive,
The millions of programs in the Blackbox dataset
have enabled dozens of contributions to the area of CS
education~\cite{bask-icer-2018}. However, these operate at the
opposite end of privacy.

@@

At the opposite end of the privacy spectrum, Go's design for \emph{transparent
telemetry} reports only counter values~\cite{transparent-telemetry}.
Unlike Roblox Studio, transparent telemetry includes no timestamps and no 
session ids.
While useful for learning about the frequency and distribution of
specific events, the lack of timestamps and ids means it cannot track
edits over time (which we use in \cref{f:error-density}).
% BEN: commented for space.
% it cannot track 
% The data is useful for learning about the frequency of specific events and the
% distribution of events across clients (e.g., how many \code{CodeTooComplex}
% errors arise? what is the average number of type errors?).
% However, the lack of timestamps and ids means that it cannot track edits
% over time (a feature that is quite useful to us, for example, in~\cref{f:error-density}).
% In~\cref{f:error-density}, for example, we see that increases to the number
% of type errors are quickly followed by decreases.
% Without timestamps, we would be unable to rule out the possibility that users
% build up a mountain of type errors and then slowly reduce them to zero.

>>>>

% TODO BEFORE SUBMISSION
% \usepackage[sort]{natbib} or otherwise manually fix cites to be in ascending order
% redo fig 5 to have better resolution

% TODO BEFORE FINAL
% - uncomment acks
% - update all refs to artifacts to no longer be in future tense

@@

\title{Privacy-Respecting Type Error Telemetry at Scale}
% lean, lite, modest, non-intrusive, private, ...
% {Millions of Type Errors} = 72M forced strict, 1/2M type error is disappointing!
%% \subtitle{Impersonal Telemetry to Measure User Experience}

@@

\newcommand{\uid}{\textsc{id}}
\newcommand{\IDE}{\textsc{ide}}
\newcommand{\code}[1]{\texttt{#1}}
\newcommand{\FILL}{\textbf{FILL}}
\newcommand{\dotscale}[1]{\scalebox{0.72}{#1}}
\newcommand{\wideas}[2]{\makebox[\widthof{#2}][l]{#1}}
\newcommand{\twoline}[2]{\parbox[s]{1.4cm}{\flushleft#1\newline#2}}
\newcommand{\chkYes}{\dotscale{\CIRCLE}}
\newcommand{\chkMaybe}{\wideas{\dotscale{\Circle}}{\chkYes}}
\newcommand{\chkNo}{\wideas{}{\chkYes}}
\newcommand{\pct}[1]{\SI{#1}{\percent}}
\newcommand{\percentile}[1]{{#1}th~percentile}
\newcommand{\ptile}[1]{\textsc{p}{#1}}
\newcommand{\modefont}[1]{\texttt{#1}}
\newcommand{\mnocheck}{\modefont{nocheck}}
\newcommand{\mnonstrict}{\modefont{nonstrict}}
\newcommand{\mstrict}{\modefont{strict}}
\newcommand{\FS}{background}
\newcommand{\zerowidth}[1]{\makebox[0pt][l]{#1}}
\newcommand{\zeroheight}[1]{\raisebox{0pt}[0pt][0pt]{\(#1\)}}
\newcommand{\nK}[1]{#1\textsc{k}}
\newcommand{\stddev}[1]{\nK{#1}}
\newcommand{\gcell}[1]{\cellcolor{green!20}#1}
\newcommand{\ycell}[1]{\cellcolor{yellow!18}#1}
\newcommand{\ocell}[1]{\cellcolor{orange!29}#1}
\newcommand{\rcell}[1]{\cellcolor{red!30}$\!\!$#1$\!\!$}
\newcommand{\gbox}[1]{\colorbox{green!20}{#1}}
\newcommand{\ybox}[1]{\colorbox{yellow!18}{#1}}
\newcommand{\obox}[1]{\colorbox{orange!29}{#1}}
\newcommand{\rbox}[1]{\colorbox{red!30}{#1}}
\newcommand{\tefsfont}[1]{\textsc{#1}}
\newcommand{\tekey}{\tefsfont{te}}
\newcommand{\fskey}{\tefsfont{bg}}
\newcommand{\QALAN}{\FILL{} Alan:}
\newcommand{\panon}{pseudonymized}
\newcommand{\symbox}[1]{{\makebox[0.8em][c]{{#1}}}}
\newcommand{\addsym}{\symbox{\(\uparrow\)}}
\newcommand{\keepsym}{\symbox{\(=\)}}
\newcommand{\dropsym}{\symbox{\(\downarrow\)}}
% \newcommand{\dsymbox}[1]{\makebox[0.8em][c]{#1}}
% \newcommand{\dropsym}{\dsymbox{\raisebox{0ex}[0.9em]{\raisebox{0.5ex}{\vrule width 0.6em height 0.6pt}}}}

@@

  \paragraph{Inquiry}
  %% What problem or question does the paper
  %% address? How has this problem or question been
  %% addressed by others (if at all)?
  Since Luau is currently being used by many creators, we want to collect data
  to improve the language and, in particular, the type system.
  The standard way to collect data is to deploy client-side telemetry; however, we cannot
  scrape personal data or proprietary information, which means we cannot
  collect source code fragments, error messages, or even filepaths.
  The research questions are thus about how to conduct telemetry that
  is not invasive and obtain insights from it about type errors.

  \paragraph{Approach}
  %% What was done that unveiled new knowledge?
  We designed and implemented a pseudonymized, randomly-sampling telemetry system for Luau.
  Telemetry records include a timestamp, a session id, a reason for sending,
  and a numeric summary of the most recent type analyses.
  This information lets us study type errors over time
  without revealing private data.
  We deployed the system in Roblox Studio during Spring 2023
  and collected over 1.5 million telemetry records from over 340,000 sessions.

@@

Creators program using the
{Luau} programming language~\cite{luau-lang.org},
an extension of {Lua~5.1~\cite{lua}}.
The main addition in Luau is a static type system that infers
types for all Luau programs on the fly, as creators modify the code.
These types are used primarily in \IDE{} tooling such as autocomplete and
in API documentation~\cite{luau-autocomplete}, and indeed, creators
may be unaware that a typechecker is analyzing their code.
However, creators can opt in to receiving type error reports and they can write
their own types to guide designs and document their intentions.

Due to the broad community of creators, the goals of the Luau type
system are rather unique~\cite{bfj-hatra-2021}.
Whereas a traditional type system focuses on compilation and memory
safety, Luau takes a lenient approach by default and lets creators
gradually~\cite{st-sfp-2006,tf-popl-2008} migrate to rigorous checks
one module at a time.
In this way, for example,
students in code camps may begin in the \mnocheck{} mode that reports only
syntax errors; hobbyists can use \mnonstrict{} mode to identify
would-be runtime errors; and developers who need to ensure software
quality can rely on \mstrict{} mode.
(No matter the mode, a {\FS{}} type analysis still drives the \IDE{} tools.)
Luau's types aim to support untyped designs, in the spirit of migratory
typing~\cite{tfffgksst-snapl-2017}, so that creators can switch modes without
needing to restructure their code in major, potentially-breaking, ways.

In this paper, we investigate methods for measuring the effectiveness
of the {Luau} type system.
The goal is to collect feedback at a large scale, with thousands
of participants maintaining real codebases.
Consequently, the measurements cannot reveal \emph{any} information about
source code, as it may contain personal data, proprietary algorithms, novel
game designs, and so on.
In comparison to prior work~(\cref{s:related}), which
with few exceptions is small in scale
or collects source code, we performed a large-scale study using \panon{}
{telemetry}.

@@

\paragraph{Contributions}
\begin{itemize}
  \item
    Design of a low-overhead telemetry system
    that reports on type errors without revealing potentially private information.

@@

{Roblox Studio} is a workbench that
combines {3D creation} tools and an Integrated
Developer Environment (\IDE{}), as seen in \cref{fig:roblox-studio}.
The \IDE{} includes an optional \emph{Script Analysis} widget that
reports syntax errors, type errors, and problems identified by
lint tools. The main editor widget can also highlight
source locations in code where reported errors occur.

\paragraph{Type Analysis Modes, Background Analysis}

@@

\begin{figure}[t]\centering
  \includegraphics[width=.35\textwidth]{img/roblox-studio.png}
  \hspace{1cm}
  \includegraphics[width=.35\textwidth]{img/roblox-studio-ide.png}
  \caption{{Roblox Studio 3D creation} tools (left) and \IDE{} (right).}
  \label{fig:roblox-studio}
\end{figure}

@@

In addition to the main type analysis, Roblox Studio takes a second pass over
every codebase with a \emph{\FS{}} analysis to infer autocomplete suggestions
and drive other \IDE{} tools.
To infer precise types,
the \FS{} analysis uses rigorous checks very similar to those of
\mstrict{} mode (but not identical: see \cref{s:strict-vs-forcedstrict})
no matter what mode the creator declared for the script.
(Internally to Roblox Studio, this analysis is called \emph{forced strict}
because it ignores the creator's choice and applies strong checks.)
When it fails to infer a type, \FS{} analysis defaults to the unknown gradual
type~\cite{st-sfp-2006}.
Creators cannot see the errors reported by \FS{} analysis.

@@

Luau modules provide reusable libraries, which may be
{imported} by other scripts.
Imports therefore form a graph in general,
though it is an error in \mstrict{} mode to create an import cycle
and edges are removed to make the graph acyclic.
%% \QALAN{}: do all modes remove the edges?

@@

Roblox Studio has a framework for reporting telemetry.
Various features in the \IDE{} use telemetry to measure their effectiveness.
The Luau open source toolchain does \emph{not} report telemetry,
as it is designed to be used in build environments such as Continuous Integration
servers, where hermetic deterministic builds are expected.

We added a subsystem to Roblox Studio's telemetry in order to study type errors
in Luau code.
This section explains the constraints our telemetry had to meet~(\cref{s:tel-limitations})
and the final design~(\cref{s:telemetry-records}).

@@

Telemetry in Roblox Studio faces three major constraints:
\begin{enumerate}

@@

\end{enumerate}
Cox~\cite{transparent-telemetry} discusses in detail the privacy implications
and tradeoffs in telemetry.

In addition, due to the architecture of Roblox Studio,
there is some information that is not available to the
typechecker:
\begin{itemize}
  \item
    Lifecycle events including as save, run, quit, and publish.\footnote{Publish events
    cannot be reported at any rate because they are public, and could be used
    to match a telemetry session to a creator.}
  \item
    GUI state: in particular, whether the Script Analysis widget is visible.
\end{itemize}
Our type error telemetry has to meet all of the above constraints.

@@

\begin{figure}[t]\centering
  \begin{tabular}{|l|l|l|l|l|l|l|}\hline
    \uid{} & Time & Mode & Reason & Size Info & Overall Counts & Edit Range Counts
    \\\toprule
    \multicolumn{6}{c}{\raisebox{.5\normalbaselineskip}[0pt][0pt]{$\underbrace{\hspace{9cm}}$}} &
    \multicolumn{1}{c}{\raisebox{.5\normalbaselineskip}[0pt][0pt]{$\underbrace{\hspace{3cm}}$}} \\[-0.5ex]
    \multicolumn{6}{c}{fixed-size} & \multicolumn{1}{c}{variable-size}
  \end{tabular}
  \caption{Structure of a telemetry record. There are 13 overall counts and up to 70 edit range counts in each record.}
  \label{f:tdata}
\end{figure}

The type error telemetry for a Roblox Studio session is gathered as a series of
records, all with the same per-session \panon{} session identifier.
This allows us to correlate telemetry across a single
session, but not between sessions.
In order to avoid swamping the telemetry servers, the client reports on a subset
of telemetry events via two levels of uniform random sampling:
\begin{itemize}
  \item
    1\% of Roblox Studio sessions generate \emph{any} type error telemetry, and
  \item
    0.5\% of all type analyses (approximately 1 in 200 keystrokes)
    in an enrolled session generate a telemetry record.
\end{itemize}
In addition to sending telemetry for randomly-selected keystrokes,
our system sends a telemetry record every time the session changes
focus from one module to another.
These records provide important context for mode switches:
if one record uses \mstrict{} mode but the next uses \mnocheck{},
then it is critical to know whether the creator downgraded modes
or simply switched to another module.

\Cref{f:tdata} shows the structure of our telemetry records:
\begin{enumerate}
  \item
    \uid{}: a \panon{} identifier for the current session.
  \item
    Time: one timestamp from the client and one from the server.
    %% timestamp explanation \cref{s:data-cleaning}
  \item
    Mode: type analysis mode of the current file: \mnocheck{},
    \mnonstrict{}, or \mstrict{}.
  \item
    Reason: a flag that explains whether this record was sent due to a
    randomly-selected keystroke or a module switch.
  \item
    Size Info: number of lines in the codebase and number of lines in the edit range.
  \item
    Overall Counts: summarizes type errors and \FS{} errors during the
    last two invocations of type analysis, and additionally reports the number
    of times type analysis hit an internal limit on problem
    size~(see~\cref{s:code-too-complex} for details).
    For each invocation and each kind of
    analysis error, there are three summary counts:
    %% grand total = 12
    \begin{itemize}
      \item total number of errors,
      \item errors in the current module, and
      \item errors in the current edit range.
    \end{itemize}
  \item
    Edit Range Counts: these list specific errors that arose in the last two
    type analyses and overlap with the current edit range.
    Since there are 35 possible errors~(\cref{s:context}), there can be up to 70 counts
    in a telemetry record.
    Exactly how to interpret this data depends on which analysis it came from:
    \begin{itemize}
      \item
        Errors in the latest type analysis are currently visible to the
        creator.
        These may have been introduced by the changes in the edit range.
      \item
        Errors from the previous type analysis were visible before the creator
        made the latest edits.
        These may have motivated some of the changes in the edit range.
    \end{itemize}
\end{enumerate}

@@

This telemetry design is admittedly coarse-grained.
For instance, it does not distinguish between edits that ignore a type
error from edits that remove one error while introducing another.
We acknowledge this and other threats in~\cref{s:threats}.
The main advantages of this telemetry are the small size of each record
and complete the lack of private information.
Furthermore, despite the limitations, this telemetry supports a variety of
inferences that we showcase in the next section.

@@

\Cref{f:records-per-hour} provides a time-ordered distribution of the data.
The x-axis is labeled with a California time zone; the thin vertical lines
mark the start of a new day.
Each day has 24 records, one per hour, represented by a blue bar
indicating the number of records generated in that hour.
Each hour has a few dozen to over 3,000 records, typically about
1,000.\footnote{We conjecture the very low bars are due to outage on
  the logging server, but need to investigate further.}
%% \FILL{} investigate one tower, what's happening?? all one session?
Telemetry ended in early April; the long tail is due to developers
updating Roblox Studio later.
% BEN: I see no basis for this; nor is this at all interesting.
%% ben: basis = it's where the peaks are
% We do not know which timezone the records originated in, but since the counts
% tend to peak midday California time it seems that the majority
% of Roblox developers are following a Western Hemisphere schedule.
The regular pattern of the peaks suggests that many creators
follow a regular schedule; these may be
professional developers.
The tall weekend (shaded) regions could be because
Roblox has a significant school-aged creator community.

@@

A close inspection of the data revealed two anomalies:
\begin{enumerate}
  \item
    Some records have identical timestamps and session \uid{}s.
    We keep the first such record and discard the rest.
    (A human would have to switch modules twice or enter several keystrokes
     within 1ms to generate two such records legitimately.)

@@

  {Among the other sessions:
        176 contain a mode upgrade,
        233 contain a mode downgrade, and
        512 contain modules with different modes.}

@@

%% mostly pure, rare mixes
Most sessions (\pct{99.82}) stick to a single analysis mode.
They never change the mode in the current module and never
switch focus to a module with a different analysis mode.
Among the other sessions, most of those (\pct{80}) switch to a
module with a different mode, about half contain at least one edit that
upgrades to a more strict mode, and about half contain a downgrade to a
less-strict mode.
There are 263 total upgrades (across all sessions) and 320 total downgrades.

@@

\subsection{Type Error Frequency}
\label{s:type-error-count}

@@

\Cref{t:type-error-count} sorts type errors by
the frequency with which they appear
in \mnonstrict{} and \mstrict{} mode.
The purpose of this table is to discover and rank all errors that
appear in the dataset; thus, it is based on both keystroke and
module-switch telemetry records, even though edit ranges are not well-defined
for the latter.
The table does not include \mnocheck{} because all its type errors are
\code{SyntaxError}s by definition.

@@

In other professional-grade \IDE{}s, such as VSCode~\cite{vscode-telemetry},
IntelliJ~\cite{intellij-telemetry}, and \code{.NET}~\cite{dotnet-telemetry}, telemetry may include all sorts of
data: error reports, source code fragments, time\-stamps, and filepaths.
Users may be able to opt out of some telemetry, but the details depend on the
license agreement.
Furthermore, at least in VSCode, \IDE{} extensions can report their own telemetry.
While this data can be invaluable for discovering bugs in production, it
must be handled with extreme care.
The Blackbox project similarly collects telemetry in broad strokes.
Every Java program submitted for compilation goes into the Blackbox database (minus
any comments above the main class) along with the compiler's response~\cite{bkmu-sigcse-2014}.
The millions of programs in the Blackbox dataset
have, however, enabled dozens of contributions to the area of CS
education~\cite{bask-icer-2018}.

@@

At the opposite end of the privacy spectrum, Go's design for \emph{transparent
telemetry} reports only counter values~\cite{transparent-telemetry}.
Unlike Roblox Studio, transparent telemetry includes no timestamps and no 
session \uid{}s.
While useful for learning about the frequency and distribution of
specific events, the lack of timestamps and \uid{}s means it cannot track
edits over time (which we use in \cref{f:error-density}).
% BEN: commented for space.
% it cannot track 
% The data is useful for learning about the frequency of specific events and the
% distribution of events across clients (e.g., how many \code{CodeTooComplex}
% errors arise? what is the average number of type errors?).
% In~\cref{f:error-density}, for example, we see that increases to the number
% of type errors are quickly followed by decreases.
% Without timestamps, we would be unable to rule out the possibility that users
% build up a mountain of type errors and then slowly reduce them to zero.
